Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-08-26 paper/ms.tex
Beef up observation section

\section{Introduction}
\the\columnwidth

\the\textwidth

% \showthe\textwidth
Approximately two-thirds of spiral disks host an elongated bar-like feature at
their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
is not difficult to explain, since stellar disks simulated in isolation almost
always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
have shown that a spherical potential, such as a stellar bulge or dark matter
halo, acts to stabilize a stellar disk against the bar instability
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

\subsection{Observations}
We note that ref.\cite{2020MNRAS.491.3655G} found that the rotation parameter
\Rot{} positively correlates with gas fraction, such that galaxies with higher
gas fractions are rotating more slowly. However, it is not obvious this is in
tension with our result since the gas fraction of galaxies correlates with
other galactic properties\cite{2009ARAA..47..159B}. Furthermore, the
measurement of pattern speeds is a delicate process still prone to large error
bars.

\subsection{Cosmological Simulations}
Barred galaxies in cosmological simulations of galaxy formation continue to be
in conflict with observations by producing bars which rotate too
slowly\cite{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}.
However, the pattern speeds of bars in both cosmological simulations and the
real universe can be affected by environmental processes not included in our
simulation -- e.g., satellite infall\cite{2011Natur.477..301P},
non-sphericity\cite{2013MNRAS.429.1949A} or rotation\cite{2013MNRAS.434.1287S,
2014ApJ...783L..18L, 2018MNRAS.476.1331C, 2019MNRAS.488.5788C} in the dark
matter halo, or perhaps even the gaseous circumgalactic medium. Naturally,
extending our present work to account for such effects is a crucial next step
in understanding the formation and evolution of galactic bars.

\section{Conclusions}
\label{sec:conclusions}
The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense\cite{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics\cite{2021MNRAS.503.2833R,
2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar\cite{2021MNRAS.500.4710C,
2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence\cite{2019MNRAS.490.4740B}. And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\cite{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time (though our work does not
alter expectations for radial mixing induced by spiral
arms\cite{2002MNRAS.336..785S}).

We found that below a certain gas fraction, bars should still be able to slow
down. Therefore, we expect barred spiral galaxies which have been gas-poor for
extended periods of time to be rotating very slowly. We therefore predict that
observations which target such galaxies (e.g., lenticular barred
galaxies\cite{2009ARAA..47..159B}) would find slowly rotating bars. There does
exist one example of a galaxy which is known to be a slow rotator -- the
low surface brightness galaxy UGC 628\cite{2009AA...499L..25C}. This galaxy
has been studied in detail by ref.\cite{2016MNRAS.463.1751C}, who note that it
indeed has a low gas fraction for galaxies of its type. We predict a general
trend that bars in gas-rich spiral galaxies should rotate quickly while bars
in gas-poor spiral galaxies should rotate slowly.

>>>>

\section{Introduction}
\label{sec:intro}
Approximately two-thirds of spiral disks host an elongated bar-like feature at
their center \citep{2000AJ....119..536E, 2007ApJ...657..790M}, including our own
Milky Way \citep{1957AJ.....62...19J, 1991ApJ...379..631B}. The ubiquity of bars
is not difficult to explain, since stellar disks simulated in isolation almost
always form bar-like structures \citep{1971ApJ...168..343H}. Several studies
have shown that a spherical potential, such as a stellar bulge or dark matter
halo, acts to stabilize a stellar disk against the bar instability
\citep{1973ApJ...186..467O, 1976AJ.....81...30H}.

@@

\subsection{Observations}
Observational estimates of the pattern speeds of bars indicate that nearly all
galaxies have $1 < \Rot < 1.4$ \citep{2011MSAIS..18...23C, 2015AA...576A.102A,
2019MNRAS.482.1733G, 2020MNRAS.491.3655G}, where $\Rot$ was defined in
Section~\ref{sec:intro} to be $\Rot\equiv \RCR/\Rb$. This observational fact has
long been in conflict with the theoretical expectation that bars should slow
down, increasing \Rot{} \citep[e.g.][]{1984MNRAS.209..729T, 1985MNRAS.213..451W,
2000ApJ...543..704D}. Explanations for this discrepancy have been given in the
past. Some have argued that perhaps the central regions of dark matter haloes
are less dense than we expected from $\Lambda\textrm{CDM}$
\citep[e.g.][]{2000ApJ...543..704D,2021AA...650L..16F}. Some have argued that
perhaps bars are recurrent, short-lived phenomena, and that all the bars we see
in the local universe are very young (CITE). Some have argued that modifications
to General Relativity ease the tension between the observed universe and
$\Lambda\textrm{CDM}$ \citep[e.g.][]{2021MNRAS.503.2833R, 2021MNRAS.508..926R}.

Because such a small gas fraction is necessary for our stabilizing mechanism to
operate ($5\%$ in our Milky Way-like disk), we argue that most galaxies host a
bar that is not slowing down. This naturally explains why most observed bars are
fast rotators. However, we acknowledge two instances of reported discrepancies
between our mechanism and observations.

First, we note that \citet{2020MNRAS.491.3655G} found that the rotation
parameter \Rot{} positively correlates with gas fraction, such that galaxies
with higher gas fractions are rotating more slowly. However, it is not obvious
this is in tension with our result since the gas fraction of galaxies correlates
with other galactic properties \citep{2009ARAA..47..159B}. Furthermore, the
measurement of pattern speeds is a delicate process still prone to large errors.

Second, the work of \citet{2021MNRAS.500.4710C} and \citet{2021MNRAS.505.2412C}
have made indirect measurements of the deceleration of the bar's pattern speed
from kinematics and chemistry. We point out that these reported measurements are
not direct measurements of the Milky Way bar's deceleration. Much like the
simulations in the present work, the simulations of these two works do not
properly account for the complicated formation process of the Galactic bar,
which may leave imprints on the present day distribution of stars in spatial,
kinematic, and chemical space. More investigation is necessary to reconcile the
present work with these two well-executed manuscripts.

\subsection{Cosmological Simulations}
Barred galaxies in cosmological simulations of galaxy formation continue to be
in conflict with observations by producing bars which rotate too slowly
\citep{2017MNRAS.469.1054A, 2019MNRAS.483.2721P, 2021AA...650L..16F}. However,
the pattern speeds of bars in both cosmological simulations and the real
universe can be affected by environmental processes not included in our
simulation -- e.g., satellite infall \citep{2011Natur.477..301P}, non-sphericity
\citep{2013MNRAS.429.1949A}, rotation in the dark matter halo
\citep{2013MNRAS.434.1287S, 2014ApJ...783L..18L, 2018MNRAS.476.1331C,
2019MNRAS.488.5788C}, or perhaps even the gaseous circumgalactic medium.
Naturally, extending our present work to account for such effects is a crucial
next step in understanding the formation and evolution of galactic bars.

\section{Conclusions}
\label{sec:conclusions}
The implications of our findings are numerous. First, we naturally explain why
nearly all observed galaxies are fast rotators without requiring the inner
regions of dark matter halos to be underdense \citep{1998ApJ...493L...5D,
2000ApJ...543..704D} or requiring new physics \citep{2021MNRAS.503.2833R,
2021MNRAS.508..926R}. Second, we show that the role of gas is of paramount
importance in studies which attempt to uncover the nature of dark matter from
its effect of slowing down the bar \citep{2021MNRAS.500.4710C,
2021MNRAS.505.2412C}. Third, we provide an explanation for how the Milky Way's
bar could be both long-lived and a fast rotator, of which there is some
observational evidence \citep{2019MNRAS.490.4740B}. And finally, we complicate
the picture of radial mixing expected to sculpt the Milky Way's disk
\citep{2012MNRAS.420..913B, 2015ApJ...808..132H}, a process which relies upon
the pattern speed of the bar to change with time (though our work does not
alter expectations for radial mixing induced by spiral
arms \citep{2002MNRAS.336..785S}).

We found that below a certain gas fraction, bars should still be able to slow
down. Therefore, we expect barred spiral galaxies which have been gas-poor for
extended periods of time to be rotating very slowly. We therefore predict that
observations which target such galaxies (e.g., lenticular barred galaxies
\citep{2009ARAA..47..159B}) would find slowly rotating bars.\footnote{This does
not mean that we predict \textit{all} gas-poor galaxies should be slowly
rotating. Indeed, they would need to be gas-deficient for several $\textrm{Gyr}$
before they would be classified as slow rotators.} There does exist one example
of a galaxy which is known to be a slow rotator -- the low surface brightness
galaxy UGC 628 \citep{2009AA...499L..25C}. This galaxy has been studied in
detail by \citet{2016MNRAS.463.1751C}, who note that it indeed has a low gas
fraction for galaxies of its type. We predict a general trend that bars in
gas-rich spiral galaxies should rotate quickly while some bars in gas-poor
spiral galaxies should rotate slowly.
